### 2017-03-25 paper/topologySkeletons.tex
wip on paper.

\subsection{ring}
\begin{center}
	\includegraphics[scale=0.75]{images/ring}
\end{center}
\begin{lstlisting}[frame=htrbl]
ring :: (ArrowLoop arr, Future fut r,
	ArrowParallel arr (i, fut r) (o, fut r) conf) =>
    conf ->
    arr (i, r) (o, r) ->
    arr [i] [o]
ring conf f =
	loop (second (rightRotate >>> lazy) >>>
    arr (uncurry zip) >>>
    parMap conf (second get >>> f >>> second put) >>>
    arr unzip)

-- from Eden, ported to Arrows:
rightRotate :: (Arrow arr) => arr [a] [a]
rightRotate = arr $ \list -> case
	list of [] -> []
			xs -> last xs : init xs
\end{lstlisting}

\subsection{torus}
\begin{center}
	\includegraphics[scale=0.75]{images/torus}
\end{center}
\begin{lstlisting}[frame=htrbl]
torus :: (ArrowLoop arr, ArrowChoice arr, ArrowApply arr,
            ArrowParallel arr (c, fut [a], fut [b]) (d, fut [a], fut [b]) conf,
            Future fut [a], Future fut [b]) =>
         conf ->
         arr (c, [a], [b]) (d, [a], [b]) ->
         arr [[c]] [[d]]
torus conf f =
	loop (second ((mapArr rightRotate >>> lazy) ***
			(arr rightRotate >>> lazy)) >>>
    arr (uncurry3 (zipWith3 lazyzip3)) >>>
    (arr length >>> arr unshuffle) &&&
        (shuffle >>> parEvalN conf (repeat (ptorus f))) >>>
    app >>> arr (map unzip3) >>> arr unzip3 >>> threetotwo)

@@

threetotwo :: (Arrow arr) => arr (a, b, c) (a, (b, c))
threetotwo = arr $ \ ~(a, b, c) -> (a, (b, c))
\end{lstlisting}

\begin{figure}[ht]
	\centering
	\includegraphics[width=0.9\textwidth]{images/torus_matrix_parrows_scale}
	\caption[without Futures]{Matrix Multiplication with a torus (Parrows)}
\end{figure}

\begin{figure}[ht]
	\centering
	\includegraphics[width=0.9\textwidth]{images/torus_matrix_eden_scale}
	\caption[with Futures]{Matrix Multiplication with a torus (Eden)}
\end{figure}

>>>>

\subsection{ring}
Eden's ring implementation allows a ring of functions to communicate over direct channels. We can rewrite its functionality easily again with the use of \code{loop}. In each loop we start by rotating the intermediary input from the nodes \code{[fut r]} with \code{second (rightRotate >>> lazy)}. We have to feed the intermediary input into \code{lazy} or the evaluation would hang. The reasoning was explained by \cite{eden_cefp}:
\begin{quotation}
Note that the list of ring inputs ringIns is the same as the list of ring outputs ringOuts rotated by one element to the right using the auxiliary function rightRotate. Thus, the program would get stuck without the lazy pattern, because the ring input will only be produced after process creation and process creation will not occur without the first input.
\end{quotation}
Next, we zip the resulting \code{([i], [fut r])} to \code{[(i, fut r)]} with \code{arr (uncurry zip)} so we can feed that into a our input arrow \code{arr (i, r) (o, r)}, which we transform into \code{arr (i, fut r) (o, fut r)} before lifting it to \code{arr [(i, fut r)] [(o, fut r)]} to get a list \code{[(o, fut r)]}. Finally we unzip this list into \code{([o], [fut r])}. Plugging this arrow \code{arr ([i], [fut r]) ([o], fut r)} into the definition of \code{loop} from earlier gives us \code{arr [i] [o]}, our ring arrow.
\begin{center}
	\includegraphics[scale=0.75]{images/ring}
\end{center}
\begin{lstlisting}[frame=htrbl]
ring :: (ArrowLoop arr, Future fut r,
	ArrowParallel arr (i, fut r) (o, fut r) conf) =>
    conf ->
    arr (i, r) (o, r) ->
    arr [i] [o]
ring conf f =
	loop (second (rightRotate >>> lazy) >>>
    arr (uncurry zip) >>>
    parMap conf (second get >>> f >>> second put) >>>
    arr unzip)
\end{lstlisting}
with \code{rightRotate}:
\begin{lstlisting}[frame=htrbl]
rightRotate :: (Arrow arr) => arr [a] [a]
rightRotate = arr $ \list -> case
	list of [] -> []
			xs -> last xs : init xs
\end{lstlisting}
and lazy:
\begin{lstlisting}[frame=htrbl]			
lazy :: (Arrow arr) => arr [a] [a]
lazy = arr (\ ~(x:xs) -> x : lazy xs
\end{lstlisting}
This combinator can be used for example to calculate the shortest paths in a graph using Warshall's algorithm. Further details on this can be found in \cite{eden_cefp}.
\subsection{torus}
\begin{center}
	\includegraphics[scale=0.75]{images/torus}
\end{center}
\begin{lstlisting}[frame=htrbl]
torus :: (ArrowLoop arr, ArrowChoice arr, ArrowApply arr,
            ArrowParallel arr (c, fut [a], fut [b]) (d, fut [a], fut [b]) conf,
            Future fut [a], Future fut [b]) =>
         conf ->
         arr (c, [a], [b]) (d, [a], [b]) ->
         arr [[c]] [[d]]
torus conf f =
	loop (second ((mapArr rightRotate >>> lazy) ***
			(arr rightRotate >>> lazy)) >>>
    arr (uncurry3 (zipWith3 lazyzip3)) >>>
    (arr length >>> arr unshuffle) &&&
        (shuffle >>> parEvalN conf (repeat (ptorus f))) >>>
    app >>> arr (map unzip3) >>> arr unzip3 >>> threetotwo)

@@

threetotwo :: (Arrow arr) => arr (a, b, c) (a, (b, c))
threetotwo = arr $ \ ~(a, b, c) -> (a, (b, c))
\end{lstlisting}
In \cite{eden_cefp}, matrix multiplication using the gentlemans algorithm was given as a example of using this skeleton. Adapting this nodefunction to our arrow API gives us:
\begin{lstlisting}[frame=htrbl]
nodefunction :: Int ->
	((Matrix, Matrix), [Matrix], [Matrix]) ->
	([Matrix], [Matrix], [Matrix])
nodefunction n ((bA, bB), rows, cols) =
	([bSum], bA:nextAs , bB:nextBs)
		where bSum =
			foldl' matAdd (matMult bA bB) (zipWith matMult nextAs nextBs)
			nextAs = take (n-1) rows
			nextBs = take (n-1) cols
\end{lstlisting}
Comparing the trace from a call using our arrow definition of the torus (fig. \ref{fig:torus_parrows_trace}) with the Eden version (fig. \ref{fig:torus_eden_trace}) shows that the behaviour of the arrow version is comparable.
\begin{figure}[ht]
	\centering
	\includegraphics[width=0.9\textwidth]{images/torus_matrix_parrows_scale}
	\caption[Matrix Multiplication with a torus (Parrows)]{Matrix Multiplication with a torus (Parrows)}
	\label{fig:torus_parrows_trace}
\end{figure}

\begin{figure}[ht]
	\centering
	\includegraphics[width=0.9\textwidth]{images/torus_matrix_eden_scale}
	\caption[Matrix Multiplication with a torus (Eden)]{Matrix Multiplication with a torus (Eden)}
	\label{fig:torus_eden_trace}
\end{figure}
